Gazing at the Robot's Head Reflects our Tendency to Anthropomorphise

### 2016-09-02 main.tex
Changed title + intro to focus on 'head gaze => anthropo'

\rightheader{Shaping the Perception of a Robot by Cognitive Priming}   % This should be the title, or a shortened version of the title if your title is long.
\leftheader{Lemaignan et al.}	  % For one or two authors, include both authors last names.  For three or more, use first author's last name et al.
\title{Shaping the Perception of a Robot by Cognitive Priming}

@@

\abstract{
What leads us to perceive robots as human-like agents? The physical appearance
of the robot and the design of its behaviours (the anthropomorphic \emph{design}
of the robot) do certainly play an important role. This paper sheds light on
another mechanism at play, less obvious because partially unconscious: the
cognitive skills that the interaction situation \emph{presupposes}.

We show that humans appear quite sensitive to this effect:
even subtle cognitive priming is sufficient to elicit significantly different
anthropomorphic attributions, and as such, this article provides new insights on
how our expectations of the robot's cognitive competency shape the bonds that we
build with the machine.

We evidence the effect with a novel methodology based on eye-tracking
and a set of stimuli that are visually identical and yet induce different
assumptions regarding the cognitive capabilities of the robot. By correlating
the resulting gaze patterns with answers to established questionnaires on
anthropomorphic attributions, we further show that fixations on the head of the robot
reflect how we anthropomorphize the machine.

@@

\begin{inparaenum}[\itshape a\upshape)]
Two nested research questions have been presented so far: \item can we
characterise the link between expectations regarding the cognitive
skills of a robot and the effective ascription of anthropomorphic features to
this robot; and \item to this end, can we develop and validate a robust behavioural and
\emph{in-the-moment} metric of anthropomorphism?  \end{inparaenum}

These two questions translate into several experimental hypotheses, whose design
was guided by the following high-level hypothesis: \emph{the more we ascribe cognitive
capabilities to an agent, the more likely we are to focus on its head (\ie the
brain) when interacting with it}. As such, this hypothesis naturally suggests an
eye-tracking-based investigation.

%Eye-tracking-based technique logically follows as an investigation mean, is a natural
%methodology to investigate this intuition. We hence design audio-visual stimuli
%that elicit different levels of cognitive ascription onto the robot while being
%identical in every other way.

Our baseline hypothesis \h{1} is that cognitive priming allows us to effectively create
stimuli that induce \emph{lower} or \emph{higher} attribution of human-likeness
to the robot. This hypothesis can be tested by asking
participants to rate the interactions depicted by the stimuli as being more
\emph{machine-like} or \emph{human-like}.

Secondly, we hypothesise that people with a higher tendency to anthropomorphise
robots will tend to physically look at robots and humans in similar ways, contrary to
people with a lower tendency to anthropomorphise (\h{2}). If this is verified,
it supports the idea that anthropomorphic attributions affect gaze patterns. The
hypothesis can be tested by correlating the participants' tendency to
anthropomorphise, measured from a standard questionnaire, with the comparative
gazing behaviour in two conditions (human task \vs robot task).

Third, we hypothesise that in the \emph{human-like} condition (which we
interpret as well as a higher ascription of cognitive skills to the robot), the participants' gaze will be directed
significantly more towards the head of the robot than in the \emph{machine-like}
condition (interpreted as a lower ascription of cognitive skills) (\h{3}). If verified, it
effectively means that the expectations held by the participants regarding the
cognitive capabilities of the robot are reflected by the eye fixations on the
head of the robot.  The hypothesis can be tested by comparing gaze fixations in
the two conditions (\emph{human-like} \vs \emph{machine-like}).

Finally, the hypothesis \h{4} proposes that manipulating cognitive expectations
affects the broader tendency to anthropomorphise robots, \ie whether observing one
robot acting in an \emph{human-like} context leads to an increased level of
anthropomorphic attributions to robots in general. The hypothesis can be
tested by looking at pre- and post-experiment self-reports on the tendency to
anthropomorphise.

If all these hypotheses are verified, the findings can be summarised as follow:
\h{1} shows that our manipulation (based on cognitive priming) does effectively
elicit lower or higher expectations in terms of the cognitive capabilities of
the robot; \h{4} shows that these expectations influence our tendency to
anthropomorphise robots in general; in parallel, \h{3} shows that these same expectations
shape the way we look at the robot by eliciting specific eye fixations on the
head of the robot; finally, \h{2} suggests that the fixation on the head
reflects human-likeliness and supports the thesis that head fixation can be used
as a behavioural measure of anthropomorphism.

@@

\section{Experimental Design}

@@

    \tikzmark{left}
    \begin{tabular}{l|p{4cm}|p{4cm}}
        & \tikzmark{top1} Machine-like cognitive priming & Human-like cognitive priming \tikzmark{top2}  \\
        \hline
        \tikzmark{left1} Robot & {\bf A}: \emph{``Pick the brown toy''} & {\bf A}: \emph{``Pick your favourite toy''} \\
                               & {\bf B}: \emph{``Point at the noise''} & {\bf B}: \emph{``Point at the crying baby''} \\
        \hline
        \tikzmark{left2} Human & {\bf A}: \emph{``Pick the brown toy''} & {\bf A}: \emph{``Pick your favourite toy''} \\ 
                               & {\bf B}: \emph{``Point at the noise''} & {\bf B}: \emph{``Point at the crying baby''}\tikzmark{bottom} \\
        \end{tabular}

@@

        \draw [decoration={brace,amplitude=0.5em},decorate,thick]
        ($(left |- bottom)+(1.8,-0.3)$) --  ($(left |- left1.north)+(1.8,0.2)$) node[midway,left=1em,rotate=90,anchor=south] {\scriptsize within subject};
    \end{tikzpicture}
    %%%

    \caption{\small The study follows a 2$\times$2 design: the \emph{robot} vs
        \emph{human} condition is within subject, while the \emph{machine-like
        cognitive priming} \vs \emph{human-like cognitive priming} is between subjects.
        Two stimuli (\emph{Picking} task, {\bf A}, and \emph{Pointing} task,
        {\bf B}) were shown to the participants, introduced by brief verbal
        commands, reproduced in the table. Figure~\ref{fig:stimuli} shows
        pictures of the four stimuli.}

@@

The audio (and audio only) of these four videos was then edited to create two
sets of stimuli: one for the \emph{machine-like cognitive priming} condition, one for
the \emph{human-like cognitive priming} condition. The audio editing consisted of 
inserting different commands to initiate the tasks (reported in
table~\ref{table:design}): in the \emph{Picking} task for instance, the object
that the agent had to pick was described either as a \emph{``brown toy''} or as
\emph{``your favourite toy''}. In the former case, the object was referred through
visible, objective features. They do not presuppose as complex cognitive capabilities
as in the latter case where the priming is expected to elicit the assumption
that the robot has tastes and preferences such as it can choose a toy as its favourite.

@@

As stated previously, the study followed a 2$\times$2 design
(table~\ref{table:design}): the \emph{machine-like} \vs \emph{human-like cognitive priming}
condition was between-subjects (30 participants performed the \emph{human-like cognitive}
tasks while 26 were in the \emph{machine-like cognitive} condition); the \emph{robot}
\vs \emph{human} condition was within-subject (the order of stimulus presentation
was counter-balanced over the participants). The study lasted in average 15
minutes per participant.

@@

We then presented the stationary eye-tracker, conducted a brief eye-tracker
calibration procedure (all participants could be successfully tracked), and the participants' gaze was recorded while they
watched the \emph{robot} and \emph{human} interaction stimuli.

@@

Participants were asked in the post-questionnaire whether the video presented a
\emph{``robot kind of task''} or a \emph{``human kind of task''} on a 5-point
scale (numerical values ranging from -2 to 2, where -2 refers to robot kind of
task). We found a significant mean difference ($M_{\text{machine-like}}=TBD (SD=TBD)$
\vs $M_{\text{human-like}}=TBD (SD=TDB)$) in this
manipulation check ($F(1,54) = 29.0, p < .001$) which validates our manipulation
and supports \h{1}: the participants that watched stimuli with a \emph{machine-like cognitive
priming} identified the task as more robot-like and participants watching
the videos with a human-like cognitive priming perceived the tasks as more human-like.

TBD: editor asks about possible ordering effect

@@

\subsection{Impact of Cognitive Priming on Changes in Anthropomorphic
Attributions}

Running an ANOVA against the change of anthropomorphic attributions
(\deltaant) over the duration of the experiment in the two cognitive
conditions, shows that \deltaant is greater for the participants who watched the
human-like tasks ($\Delta_{\text{machine-like}}=TBD (SD=TBD)$
\vs $\Delta_{\text{human-like}}=TBD (SD=TDB)$, $F(1,54) = 6.54, p < .05$\fixme{put the exact
p value here}). In other words, after being exposed to short videos of a robot
performing simple tasks, participants primed with a richer cognitive
expectations regarding the robot report a significantly higher tendency to anthropomorphise robots
\emph{in general} (as the questionnaire on anthropomorphic attributions
explicitly referred to \emph{robots in general}).

This result supports \h{4}: cognitive priming impacts the
attribution of anthropomorphic features to robots \emph{beyond the specific Nao robot
shown in the stimuli}.  It appears that participants transfer to some extent the
cognitive capabilities that they ascribe to the Nao robot in the
\emph{human-like} cognitive condition to ``robots'' in general (we can not, however,
conclude on any possible \emph{lasting} anthropomorphic ascriptions to robots as we
measure the anthropomorphic perception immediately after the experiment).

@@

\section{Conclusion}

@@

The second contribution is a methodological one: we present in this article
a novel technique to assess anthropomorphic projections that relies on a
biometric measurement (eye-tracking) to compare gaze fixation durations on
the face of the robot. We cross-validate this new metric with existing,
established, questionnaires. Compared to current techniques (post-hoc
annotations of videos and questionnaires), this new approach is objective,
is less impacted by experimental biases (like the \emph{observer effect},
where the behaviour of the participant is impacted by the fact he/she knows that
he/she is observed) and take place during the interaction itself
(\emph{in-the-moment} measurement).

>>>>

\rightheader{Gaing at the Robot's Head Reflects our Tendency to Anthropomorphise}   % This should be the title, or a shortened version of the title if your title is long.
\leftheader{Lemaignan et al.}	  % For one or two authors, include both authors last names.  For three or more, use first author's last name et al.
\title{Gazing at the Robot's Head Reflects our Tendency to Anthropomorphise}

@@

\abstract{

Could the way we look at robots betray our anthropomorphic ascriptions to the
machine? We show in the article that indeed, the duration of gaze fixations on
the head of a humanoid robot reflects one's tendency to anthropomorphise the
robot.  We reach this conclusion by building on an novel experimental protocol
involving cognitive priming: participants are shown a set of stimuli that are
visually identical and yet induce different assumptions regarding the cognitive
capabilities of the robot. By correlating the resulting gaze patterns with
answers to established questionnaires on anthropomorphic attributions, we show
that fixations on the head of the robot reflect how we anthropomorphize the
machine. Besides, by comparing this gazing behaviour to a human baseline, we further
show that people with a generally higher tendency to anthropomorphise exhibit
similar gaze patterns when they look either at humans or robots.

@@

As suggested in the previous section, we aim at verifying the following research
hypothesis \h{}: the level of anthropomorphism ascribed by a human to a robot can be
effectively assessed by measuring the human's gaze fixations on the robot's head.

Our methodology involves showing participants video stimuli that elicit
different levels of anthropomorphism, and measuring gaze fixations over the
course of the video clips. To be suitable for an eye-tracking experiment, the
video stimuli must be visually identical between conditions, and therefore
we have to rely on non-visual cues to elicit different anthropomorphic
projections. We choose to use cognitive priming with audio cues, as detailed in
the next section.

This approach allows us to decompose the hypothesis \h{} into two simpler
hypotheses \h{1} and \h{2}: we want first to verify that cognitive priming
enables us to effectively manipulate the attribution of human-likeness to robots
(\h{1}).  We need next to verify that in the case of higher anthropomorphic
attributions, the participants' gaze will be directed significantly more towards
the head of the robot than in the case of low anthropomorphic attributions
(\h{2}).

If verified, these hypotheses would allow us to conclude that in a given
human-robot interaction situation, gaze fixations on a robot's head reflects the
ascription of anthropomorphic features to the robot. This would suggest that a
relatively simple biometric and in-the-moment measure (eye-tracking) can
effectively inform us about occurring anthropomorphic projections.

\vspace{1em}

Besides, our experimental design allows to examine two additional hypotheses:

\begin{itemize}
    \item people who generally tend to anthropomorphise exhibit gazing
        behaviours that are similar when looking either at humans or robots, and
        reciprocally (\h{3}). If this is verified, this would confirm that our
        gazing behaviour towards agents is impacted by our tendency to
        anthropomorphise.

    \item the mere observation of a robot acting in an \emph{human-like}
        situation suffice to increase the level of
        anthropomorphic attributions to robots \emph{in general} (\h{4}).
\end{itemize}

%If all these hypotheses are verified, the findings can be summarised as follow:
%\h{1} shows that our manipulation (based on cognitive priming) does effectively
%elicit lower or higher expectations in terms of the cognitive capabilities of
%the robot; \h{4} shows that these expectations influence our tendency to
%anthropomorphise robots in general; in parallel, \h{3} shows that these same expectations
%shape the way we look at the robot by eliciting specific eye fixations on the
%head of the robot; finally, \h{2} suggests that the fixation on the head
%reflects human-likeliness and supports the thesis that head fixation can be used
%as a behavioural measure of anthropomorphism.

@@

\section{Experimental Design}
\label{sec:design}

@@

    \tikzmark{left}
    \begin{tabular}{l|p{4.3cm}|p{4.3cm}}
        & \tikzmark{top1} Machine-like cognitive priming & Human-like cognitive priming \tikzmark{top2}  \\
        \hline
        \tikzmark{left1} Robot & {\bf A}: \emph{``Pick the brown toy''} & {\bf A}: \emph{``Pick your favourite toy''} \\
                               & {\bf B}: \emph{``Point at the noise''} & {\bf B}: \emph{``Point at the crying baby''} \\
        \hline
        \tikzmark{left2} Human & {\bf A}: \emph{``Pick the brown toy''} & {\bf A}: \emph{``Pick your favourite toy''} \\ 
                               & {\bf B}: \emph{``Point at the noise''} & {\bf B}: \emph{``Point at the crying baby''}\tikzmark{bottom} \\
        \end{tabular}

@@

        \draw [decoration={brace,amplitude=0.5em},decorate,thick]
        ($(left |- bottom)+(1.4,-0.3)$) --  ($(left |- left1.north)+(1.4,0.2)$) node[midway,left=1em,rotate=90,anchor=south] {\scriptsize within subject};
    \end{tikzpicture}
    %%%

    \caption{\small The study follows a 2$\times$2 design: the \emph{robot} vs
        \emph{human} condition is within subject, while the \emph{machine-like
        cognitive priming} \vs \emph{human-like cognitive priming} is between subjects.
        Two stimuli (\emph{Picking} task {\bf A}, and \emph{Pointing} task
        {\bf B}) were shown to the participants, introduced by brief verbal
        commands, reproduced in the table. Figure~\ref{fig:stimuli} shows
        pictures of the four stimuli.}

@@

The audio (and audio only) of these four videos was then edited to create two
sets of stimuli: one for the \emph{machine-like cognitive priming} condition, one for
the \emph{human-like cognitive priming} condition. The audio editing consisted of 
inserting different commands to initiate the tasks (reported in
figure~\ref{table:design}): in the \emph{Picking} task for instance, the object
that the agent had to pick was described either as a \emph{``brown toy''} or as
\emph{``your favourite toy''}. In the former case, the object was referred through
visible, objective features. They do not presuppose as complex cognitive capabilities
as in the latter case where the priming is expected to elicit the assumption
that the robot has tastes and preferences such as it can choose a toy as its favourite.

@@

As stated previously, the study followed a 2$\times$2 design
(table~\ref{table:design}): the \emph{machine-like} \vs \emph{human-like
cognitive priming} condition was between-subjects. 30 participants performed the
\emph{human-like cognitive} tasks while 26 were in the \emph{machine-like
cognitive} condition. The \emph{robot} \vs \emph{human} condition was
within-subject. The study lasted in average 15 minutes per participant.

@@

We then presented the stationary eye-tracker, conducted a brief eye-tracker
calibration procedure (all participants could be successfully tracked), and the
participants' gaze was recorded while they watched the \emph{robot} and
\emph{human} interaction stimuli. The order of stimulus presentation
was counter-balanced over the participants.

@@

Participants were asked in the post-questionnaire whether the video presented a
\emph{``robot kind of task''} or a \emph{``human kind of task''} on a 5-point
scale (numerical values ranging from -2 to 2, where -2 refers to robot kind of
task). We found a significant mean difference ($M_{\text{machine-like}}=TBD (SD=TBD)$
\vs $M_{\text{human-like}}=TBD (SD=TDB)$) in this
manipulation check ($F(1,54) = 29.0, p < .001$): the participants who watched
stimuli in the \emph{machine-like priming} condition identified the tasks as more
robot-like, while participants who watched the videos with a \emph{human-like
priming} perceived the tasks as more human-like.

Note that, expectedly, we found an order effect on the presentation of the robot
and human stimuli (our within-subject variable): if the stimuli shown last
involved a human, the participant reported a more human-like kind of tasks
(ANOVA between the models with condition only and condition and order: $F(1,51)
= 11.15, p < .001$, no interaction). This was accounted for by counter-balancing
the presentation order.

This first result supports \h{1}: cognitive priming enables us to
effectively elicit different levels of human-likeliness attribution.
\h{1} is however only partially supported as it specifically hypothesised that cognitive
priming can manipulate anthropomorphic attributions to the \emph{robot} itself:
the results of the manipulation check only indicate that the priming impacted
the participants' anthropomorphic perception of the \emph{interaction
situation} as a whole.

\subsection{Impact of Cognitive Priming on Anthropomorphic Attributions}

Running an ANOVA against the change of anthropomorphic attributions
\deltaant before and after the experiment shows that \deltaant is greater for the participants who watched the
human-like tasks ($\Delta_{\text{machine-like}}=TBD (SD=TBD)$
\vs $\Delta_{\text{human-like}}=TBD (SD=TDB)$, $F(1,54) = 6.54, p < .05$\fixme{put the exact
p value here}). In other words, after being exposed to short videos of a robot
performing simple tasks, participants primed with richer cognitive
expectations regarding the robot report a significantly higher tendency to anthropomorphise robots
\emph{in general} (as the questionnaire on anthropomorphic attributions
explicitly referred to \emph{robots in general}).

This result shows that the manipulation of the cognitive requirements of our
tasks does indeed alter the participants' perception of the robots themselves.
Combined with the previous result, we can conclude that \h{1} is indeed
supported: by the mean of cognitive priming, we can manipulate the
human-likeness attributions to robots.

This result also supports \h{4}: cognitive priming impacts the
attribution of anthropomorphic features to robots beyond the specific Nao robot
shown in the stimuli. It appears that participants transfer to some extent the
cognitive capabilities that they ascribe to the Nao robot in the
\emph{human-like} cognitive condition to robots in general. We can not, however,
conclude on any possible \emph{lasting} anthropomorphic ascriptions to robots as we
measure the anthropomorphic perception immediately after the experiment.

@@

@@

\section{Conclusion}
\label{conclusion}

@@

The second contribution is a methodological one: we present in this article
a novel technique to assess anthropomorphic projections that relies on a
biometric measure (eye-tracking) to compare gaze fixation durations on
the face of the robot. We cross-validate this new metric with existing,
established, questionnaires. Compared to current techniques (post-hoc
annotations of videos and questionnaires), this new approach is objective,
is less impacted by experimental biases (like the \emph{observer effect},
where the behaviour of the participant is impacted by the fact he/she knows that
he/she is observed) and take place during the interaction itself
(\emph{in-the-moment} measurement).
